\section{Introduction}
\subsection{Motivation}
In recent years, the increasing reliance on centralized communication infrastructures has exposed significant vulnerabilities during emergencies and connectivity outages. Natural disasters, political unrest, and network restrictions often render traditional internet-based communication useless, leaving people isolated and unable to coordinate or request help.  
Furthermore, growing privacy concerns highlight the risks of centralized servers that can collect, monitor, or expose user data.  

This project is motivated by the need for a secure, privacy-preserving, and infrastructure-independent communication system. By leveraging Bluetooth Mesh technology and end-to-end encryption, the proposed system enables direct, resilient, and private communication between users, even in the absence of mobile networks or Wi-Fi.


\subsection{Objectives}
The primary objectives of this project are as follows:

\begin{itemize}
    \item \textbf{To design and develop} a decentralized peer-to-peer messaging application using Bluetooth Mesh communication.
    \item \textbf{To ensure security and privacy} through end-to-end encryption, authentication, and integrity protection.
    \item \textbf{To support offline communication} in environments where internet or cellular networks are unavailable or unreliable.
    \item \textbf{To optimize performance and battery efficiency} by using Bluetooth Low Energy (BLE) for background operation.
    \item \textbf{To provide a user-friendly interface} that allows non-technical users to communicate easily without complex setup.
    \item \textbf{To evaluate system scalability and reliability} across different real-world use cases such as disaster zones, remote communities, and crowded public events.
\end{itemize}
\subsection{Scope}
The scope of this project includes the design, implementation, and testing of a Bluetooth Mesh-based communication system that enables short- to medium-range peer-to-peer messaging. The project focuses on local device-to-device networking without reliance on internet or mobile infrastructure.  

Key features within the project scope include:
\begin{itemize}
    \item Bluetooth Mesh message routing, node discovery, and relay functionality.
    \item End-to-end encryption for all transmitted data.
    \item Text and small media (e.g., image) message exchange.
    \item Android-based implementation using Kotlin and BLE APIs.
\end{itemize}

Out of scope for this project:
\begin{itemize}
    \item Integration with cloud or server-based systems.
    \item Long-range or cross-network message relaying via the internet.
    \item Support for iOS or other operating systems in the current version (planned for future expansion).
\end{itemize}

The project will thus deliver a fully functional Android prototype demonstrating secure, decentralized, and offline communication over Bluetooth Mesh networks.
\subsection{Significance of the Project}

\subsubsection{Significance from the Practical Perspectives}

\subsubsection{Significance from the Scientific Perspectives}

\subsection{Report structure}