\section{Conclusion}
This project presents the design and development of a decentralized peer-to-peer messaging application operating over Bluetooth mesh networks. By leveraging Bluetooth Low Energy technology and implementing end-to-end encryption, the system addresses a critical gap in communication infrastructure by enabling secure and reliable messaging in environments where traditional internet or cellular networks are unavailable or unreliable. The application demonstrates practical solutions for scenarios such as natural disasters, remote areas, large public gatherings, and situations in which communication infrastructure is deliberately restricted. \\
The system architecture adopts a layered monolithic design that balances modularity, maintainability, and performance constraints inherent to mobile devices. Through careful selection of cryptographic primitives, including Curve25519 for key exchange, ChaCha20-Poly1305 for authenticated encryption, and the Noise Protocol Framework for secure channel establishment, the application supports three distinct security levels to accommodate different usage scenarios and trust requirements. The implementation of the Hybrid Multi-hop Forwarding Over Routing (HMFOR) algorithm enables efficient message propagation across the mesh network while minimizing control overhead and dynamically adapting to topology changes. \\
The three-tier identity management system consists of usernames, dynamic routing identifiers, and static public keys. This approach successfully reconciles the competing demands of usability, privacy, and cryptographic assurance. By deriving Mesh Protocol Addresses dynamically from static keys using HKDF with time-based epochs, the system achieves forward privacy and mitigates device tracking while preserving routing functionality. \\
The protocol design includes system packets for network maintenance and user packets for application-level communication, providing a comprehensive framework for reliable and secure data exchange in a fully decentralized environment. Implementation using Kotlin and Android Studio, together with SQLite for local data persistence and native BLE APIs for connectivity, demonstrates the feasibility of deploying sophisticated mesh networking capabilities on standard consumer devices without requiring specialized hardware. \\
The user interface design prioritizes simplicity and accessibility, allowing non-technical users to communicate effectively in challenging environments without complex configuration procedures. However, several challenges and limitations remain. Bluetooth’s inherent range constraints, limited bandwidth, and susceptibility to interference impose fundamental restrictions on network scale and message throughput. The current implementation does not fully address scenarios involving very large networks beyond 100 to 500 nodes, prolonged device mobility, or advanced adversarial threats such as Sybil attacks or targeted jamming. In addition, features such as voice communication, large file transfers, and cross-platform support remain areas for future development. \\
Looking ahead, the next development phase will focus on completing the implementation roadmap outlined in the project plan. This includes comprehensive security testing, performance optimization under real-world conditions, and evaluation of a hybrid centralized-decentralized topology variant. Further research is required to assess system behavior across diverse deployment scenarios and to refine routing strategies based on empirical performance data. The integration of additional out-of-band verification methods and the exploration of energy-efficient duty cycling mechanisms are expected to enhance both security guarantees and battery longevity. \\
In conclusion, this project demonstrates that infrastructure-independent and secure peer-to-peer communication over Bluetooth mesh networks is both technically feasible and practically valuable. By combining modern cryptographic techniques with adaptive networking protocols, the system provides a resilient communication platform that enables individuals to maintain connectivity when conventional infrastructure fails. The insights gained from this work contribute to a broader understanding of decentralized communication systems and provide a foundation for future innovations in off-grid and privacy-preserving messaging technologies. \\